---
title: "Eidolon_phenology_modis_modeled"
output: html_document
editor_options: 
  chunk_output_type: inline
---
# load libraries
```{r setup, echo=FALSE}
# packrat::init()
# basic toolkit
library(ggplot2)
library(ggformula)
library(dplyr) 
library(tidyverse) # piping
library(scales) 
library(viridisLite) # nice colors
library(ggpubr) # nice plotting
library(readxl) # read xls
library(janitor) # clean column names
library(zoo) # interpolation
library(raster)

# time series
library(lubridate)

# math
require(pracma) # area under the curve

# Mapping
library(rnaturalearth)
library(rnaturalearthdata)

# Species occurrence data
library(rgbif)
library(rinat)

# modeling
library(mgcv)
library(lme4)
library(MuMIn)
# library(tidymodels)
library(vip)


source("../src/IRG_functions.R")
```

# load counts
```{r}
load("../../../../Dropbox/MPI/Eidolon/Greenwave/rdata/colony_count.RData")
```


# load GEE data
```{r load GEE data}

EVI <- read.csv("../../../../Dropbox/GreenWave/EidolonColonies/rawdata/rl_MOD13Q1_EVI_2000_2020_ptsB_EidolonColonies_QCOP1_buf50km.csv")
MEVI1 <- read.csv("../../../../Dropbox/GreenWave/EidolonColonies/rawdata/MODIS_Phen_buf67/p_MCD12Q2_Peak_1_2000_2020_ptsB.csv")
table(MEVI1$geeID)
MEVI2 <- read.csv("../../../../Dropbox/GreenWave/EidolonColonies/rawdata/MODIS_Phen_buf67/p_MCD12Q2_Peak_2_2000_2020_ptsB.csv")
MIRG1 <- read.csv("../../../../Dropbox/GreenWave/EidolonColonies/rawdata/MODIS_Phen_buf67/p_MCD12Q2_MidGreenup_1_2000_2020_ptsB.csv")
MIRG2 <- read.csv("../../../../Dropbox/GreenWave/EidolonColonies/rawdata/MODIS_Phen_buf67/p_MCD12Q2_MidGreenup_2_2000_2020_ptsB.csv")
MGRD1 <- read.csv("../../../../Dropbox/GreenWave/EidolonColonies/rawdata/MODIS_Phen_buf67/p_MCD12Q2_MidGreendown_1_2000_2020_ptsB.csv")
MGRD2 <- read.csv("../../../../Dropbox/GreenWave/EidolonColonies/rawdata/MODIS_Phen_buf67/p_MCD12Q2_MidGreendown_2_2000_2020_ptsB.csv")
numcycles <- read.csv("../../../../Dropbox/GreenWave/EidolonColonies/rawdata/MODIS_Phen_buf67/p_MCD12Q2_NumCycles_2000_2020_ptsB.csv")
QA1 <- read.csv("../../../../Dropbox/GreenWave/EidolonColonies/rawdata/MODIS_Phen_buf67/p_MCD12Q2_QA_Overall_1_2000_2020_ptsB.csv")
QA2 <- read.csv("../../../../Dropbox/GreenWave/EidolonColonies/rawdata/MODIS_Phen_buf67/p_MCD12Q2_QA_Overall_2_2000_2020_ptsB.csv")
MEVImin1 <- read.csv("../../../../Dropbox/GreenWave/EidolonColonies/rawdata/MODIS_Phen_buf67/p_MCD12Q2_EVI_Minimum_1_2000_2020_ptsB.csv")
MEVImin2 <- read.csv("../../../../Dropbox/GreenWave/EidolonColonies/rawdata/MODIS_Phen_buf67/p_MCD12Q2_EVI_Minimum_2_2000_2020_ptsB.csv")
MEVIamp1 <- read.csv("../../../../Dropbox/GreenWave/EidolonColonies/rawdata/MODIS_Phen_buf67/p_MCD12Q2_EVI_Amplitude_1_2000_2020_ptsB.csv")
MEVIamp2 <- read.csv("../../../../Dropbox/GreenWave/EidolonColonies/rawdata/MODIS_Phen_buf67/p_MCD12Q2_EVI_Amplitude_2_2000_2020_ptsB.csv")
MEVIarea1 <- read.csv("../../../../Dropbox/GreenWave/EidolonColonies/rawdata/MODIS_Phen_buf67/p_MCD12Q2_EVI_Area_1_2000_2020_ptsB.csv")
MEVIarea2 <- read.csv("../../../../Dropbox/GreenWave/EidolonColonies/rawdata/MODIS_Phen_buf67/p_MCD12Q2_EVI_Area_2_2000_2020_ptsB.csv")

precip <- read.csv("../../../../Dropbox/MPI/Eidolon/Greenwave/data/GEE/ee_precip_chirps_colonies_2000_2020.csv")
names(precip) <- c("time", 1:17)
precip$time <- dmy(precip$time)
```

## join data
```{r join data}
MEVI1$amp <- MEVIamp1$mean
MEVI2$amp <- MEVIamp2$mean
MEVI1$min <- MEVImin1$mean
MEVI2$min <- MEVImin2$mean
MEVI1$max <- MEVI1$amp + MEVI1$min
MEVI2$max <- MEVI2$amp + MEVI2$min
MEVI1$area <- MEVIarea1$mean
MEVI2$area <- MEVIarea2$mean
MEVI1$IRG <- MIRG1$mean
MEVI2$IRG <- MIRG2$mean
MEVI1$GRD <- MGRD1$mean
MEVI2$GRD <- MGRD2$mean
MEVI1$peak <- 1
MEVI2$peak <- 2
MEVI1$QA <- QA1$mean
MEVI2$QA <- QA2$mean

MEVI <- rbind(MEVI1, MEVI2)

MEVI$date_EVI <- ymd("1970-01-01")+MEVI$mean
MEVI$yday_EVI <- yday(MEVI$date_EVI)
MEVI$date_IRG <- ymd("1970-01-01")+MEVI$IRG
MEVI$yday_IRG <- yday(MEVI$date_IRG)
MEVI$date_GRD <- ymd("1970-01-01")+MEVI$GRD
MEVI$yday_GRD <- yday(MEVI$date_GRD)

MEVI <- MEVI[order(MEVI$date_EVI),]
# MEVI$lat <- unique(colonies$Lat)

for(i in 1:nrow(MEVI)){
  MEVI$lat[i] <- unique(colonies$Lat[MEVI$geeID[i] == colonies$geeID])
  MEVI$Location[i] <- unique(colonies$Location[MEVI$geeID[i] == colonies$geeID])
}

summary(MEVI)

EVI$time <- ymd_hms(EVI$startDate)
```

# add EVI, IRG, and precipitation to all counts
```{r colony cycles}
IDs <- unique(reg_colonies$geeID)
i = 14
layout(1)
for(i in 1:length(IDs)){
  c <- reg_colonies[reg_colonies$geeID == IDs[i],]
  e <- EVI[EVI$geeID == IDs[i],]
  p <- precip[,c(1,(i+1))]
  names(p) <- c("time", "mean")
  # layout(rbind(1,2))

  with(c, plot(month, ratio, type = "o", 
               pch = 16, col = as.numeric(is.na(Observer))+1,
               ylim = c(0,1),
               main = paste0(Location[1],", ", Country[1] , "; ID = ", geeID[1])))
  with(e, points(as.Date(time), mean/max(mean), type = "o", col = 3, pch = 16))
  with(p, points(as.Date(time), mean/max(mean), type = "o", col = 4, pch = 16))
  with(c, points(month, ratio, type = "o", 
               pch = 16, col = as.numeric(is.na(Observer))+1))
}

```
# add IRG to EVI
```{r}
IDs
EVI$EVI <- NA
EVI$IRG <- NA
EVI$Lat <- NA
i = 1
for(i in 1:length(IDs)){
  idx <- which(EVI$geeID == i)
  e <- EVI[idx,]
  time <- e$time
  evi <- e$mean
  
  spl <- smooth.spline(x = time, y = evi)
  pred <- predict(spl)
  EVI$EVI[idx] <- pred$y
  # get derivative of EVI -> IRG
  pred.prime <- predict(spl, deriv=1)
  plot(pred.prime, type = "l")
  # pred.prime$y_scale <- rescale(pred.prime$y, c(0,1))
  # plot(pred.prime$x, pred.prime$y_scale, type = "l")
  
  EVI$IRG[idx] <- pred.prime$y
  
  EVI$Lat[idx] <- colonies$Lat[which(colonies$geeID == i)[1]]
}
```


# ggplot
```{r}
ggplot(EVI, aes(x = month(time)))+
  geom_smooth(col = 3, aes(y = EVI/max(EVI), se = FALSE, group = geeID))+
  geom_smooth(aes(y = IRG/max(IRG), group = geeID), se = FALSE, col = 2)+
  geom_smooth(data = reg_colonies, se = FALSE, 
              aes(x = month(month), y = ratio, group = geeID), col = 1)+
  # coord_cartesian(xlim=c(0,12))+
  xlab("ratio")+ylab("month")+theme_classic()+
  facet_wrap(~round(Lat,2), ncol = 6)

```


# Find corresponding EVI and precipitation data for all colony data
```{r}
IDs <- 1:max(colonies$geeID)
rs <- data.frame()
i = 12
for(i in 1:length(IDs)){
  ID <- i
  e <- EVI[EVI$geeID == ID,]
  
  p <- precip[,c(1,ID+1)]
  names(p) <- c("time", "precip")
  p$time <- ymd_hms(paste0(p$time, " 12:00:00"))
  
  c_idx <- which(colonies$geeID == ID)
  colony <- colonies[c_idx,]
  colony$time <- ymd(colony$date)
    
  par(mar = c(1,4,1,1), oma = c(1,1,1,1), xpd=FALSE)
  # layout(rbind(1,2))
  layout(1)
  colony$time-1
  with(colony, plot(time, ratio, type = "o", lwd = 2,
       xlim = c(min(time)-365, max(time)+365)))
  lines(as.Date(e$time), e$mean, col = 3, lwd = 2,)
  lines(x = as.Date(p$time), y = p$precip/max(p$precip), lwd = 2, col = 4)
  
  mevi1 <- MEVI[MEVI$geeID == ID & MEVI$peak == 1,]
  mevi2 <- MEVI[MEVI$geeID == ID & MEVI$peak == 2,]
  
  points(x = mevi1$date_EVI, col = "green",
         y = rep(1, nrow(mevi1)), pch = 16)
  points(x = mevi2$date_EVI, col = "darkgreen",
         y = rep(1, nrow(mevi2)), pch = 16)
  
  points(x = mevi1$date_IRG, col = "red",
         y = rep(1, nrow(mevi1)), pch = 16)
  points(x = mevi2$date_IRG, col = "darkred",
         y = rep(1, nrow(mevi2)), pch = 16)
  
  points(x = mevi1$date_GRD, col = "blue",
         y = rep(1, nrow(mevi1)), pch = 16)
  points(x = mevi2$date_GRD, col = "darkblue",
         y = rep(1, nrow(mevi2)), pch = 16)
  
  legend("topright", legend = c("IRG1","IRG2","EVI1", "EVI2", "GRD1", "GRD2"), 
         col = c("red", "darkred", "green", "darkgreen", "blue", "darkblue"), pch = 16)
  
  dl <- geosphere::daylength(colony$Lat[1], seq(1,331, by = 30))
  dayl <- data.frame(date = seq.Date(as.Date("2000-01-01"), 
                                     as.Date("2019-12-01"), 
                                     by = "month"), 
                     daylength = rep(dl, 20))
  dayl$geeID <- ID
  dayl$year <- year(dayl$date)
  lines(dayl$date, rescale(dayl$daylength), 
        col = "gold", lwd = 2)
  
    summary(p)
  p$year <- year(p$time)
  p$geeID <- ID
  p_sum <- p %>% group_by(year, geeID) %>% 
    summarise(precip_spline = time[which.max(precip)])

  MEVI$year <- year(MEVI$date_EVI)
  m_sum1 <- MEVI[MEVI$geeID == ID,] %>% group_by(geeID, year) %>% 
    summarise(EVI1_model = date_EVI[which(peak == 1)],
              EVI_model = date_EVI[which.max(amp)],
              IRG1_model = date_IRG[which(peak == 1)],
              IRG_model = date_IRG[which.max(amp)],
              GRD1_model = date_GRD[which(peak == 1)],
              GRD_model = date_GRD[which.max(amp)],
              amp_MEVI = amp[which.max(amp)],
              min_MEVI = min[which.max(amp)],
              max_MEVI = max[which.max(amp)])
  
  m_sum2 <- MEVI[MEVI$geeID == ID,] %>% group_by(geeID, year) %>% 
    summarise(EVI2_model = date_EVI[which(peak == 2)],
              IRG2_model = date_IRG[which(peak == 2)],
              GRD2_model = date_IRG[which(peak == 2)])
  m_sum <- full_join(m_sum1, m_sum2)
  
  e$year <- year(e$time)
  e_sum <- e %>% group_by(geeID, year) %>% 
    summarise(max_EVI = max(EVI, na.rm = TRUE),
              min_EVI = min(EVI))
  # dayl$daylength %>% plot(type = "l")
  # dl_sum <- dayl %>% group_by(geeID, year) %>%
  #   summarise(daylength_month = date[which.max(daylength)],
  #             daylength = daylength[which.max(daylength)])

  # plot(dayl$daylength)
  C <- full_join(p_sum, m_sum)
  C <- full_join(C, e_sum)
  rs <- rbind(rs, C)
}
summary(rs)
save(colonies, rs, file = "../../../../Dropbox/MPI/Eidolon/Greenwave/rdata/colonies_MEVI.Rdata")

# (rs$daylength) %>% plot
```  

```{r}
plot(rs$max_EVI, rs$max_MEVI)
plot(rs$min_EVI, rs$min_MEVI)
rs$amp_EVI <- rs$max_EVI - rs$min_EVI
plot(rs$amp_MEVI, rs$amp_EVI)

```



# Check peak colony month
## find colony peaks
```{r}
Peaks <- data.frame()
IDs <- unique(colonies$geeID)
i = 14
for(i in 1:length(IDs)){
  rc <- colonies[colonies$geeID == IDs[i],]
  rc <- rc[order(rc$date),]
  # find peaks
  peak_time <- rc$date[findpeaks(rc$ratio)]
  
  # add any missed peaks at ratio of .6
  peak_time <- unique(c(peak_time, na.omit(rc$date[rc$ratio > 0.6])))
  
  peaks <- rc[which(rc$date %in% peak_time),]
  
  # plot peaks
  with(rc, plot(date, ratio, type = "o", pch = 16,
                col = as.numeric(is.na(Observer))+1,
                main = rc$Location[1]))
  abline(v = peaks$date, lty = 2) 
  
  # remove peaks with ratios below 0.4
  peaks <- peaks[which(peaks$ratio > 0.4),]
  
  
  # remove peaks that are within 6 months
  while(any(diff(peaks$date) < 180)){
    
    too_close <- which(diff(peaks$date) < 180)[1]
    
    # if sequential peaks with ratio = 1, then take the middle
    if(all(peaks$ratio[too_close:(too_close+1)] == 1)){
      idx <- which((peaks$date[too_close:(too_close+10)] - peaks$date[too_close]) < 180) 
      tmp <- rle(peaks$ratio[too_close:(too_close+(length(idx)-1))])
      ctmp <- rle(peaks$Count[too_close:(too_close+(length(idx)-1))])
      if(tmp$lengths[1] == ctmp$lengths[1]){
        # even
        if((tmp$lengths[1] %% 2) == 0){
          keep <- round(median(1:tmp$lengths[1]),0)
          kick <- (1:tmp$lengths[1])[-keep]
          peaks <- peaks[-(too_close+(kick-1)),]
        }
        # odd
        if((tmp$lengths[1] %% 2) != 0){
          keep <- median(1:tmp$lengths[1])
          kick <- (1:tmp$lengths[1])[-keep]
          peaks <- peaks[-(too_close+(kick-1)),]
        }
      }
        
    }
    # update too_close
    too_close <- which(diff(peaks$date) < 180)[1]
    
    # keep peak with higher ratio
    if(any(peaks$ratio[too_close:(too_close+1)] != 1))
    kick <- which.min(peaks$ratio[too_close:(too_close+1)])
    peaks <- peaks[-(too_close+kick-1),]
  }
  abline(v = peaks$date, lwd = 2, col = 4) 
  Peaks <- rbind(Peaks, peaks)
}
Peaks$month <- round_date(Peaks$date, unit = "month")

```

# plot colony and RS peaks 
```{r}
i = 12
for(i in 1:length(IDs)){
  p <- Peaks[which(Peaks$geeID == IDs[i]),]
  r <- rs[which(rs$geeID == IDs[i]),]
  # layout(c(1,2))
  
  plot(r$year, month(r$IRG_model), type = "o", pch = 16,
       ylab = "month", xlab = "", col = 2,
       ylim = c(0,12), main = paste(p$Location[1], "model"))
  # abline(v = 2000:2020, lty = 2)
  points(r$year, month(r$EVI_model), type ="o", pch = 16, col = 3)
  points(r$year, month(r$precip_spline), type = "o", pch = 16, col = 4)
  # points(r$year, month(r$daylength_month), type = "o", pch = 16, col = "gold")
  points(x = year(p$month), y = month(p$month), col = 1, pch = 16, cex = 2)
  
  
  legend("bottomleft", legend = c("EVI", "IRG", "Precip", "Colony"), pch = 16, col = c(3,2,4,1))
  # plot(r$year, month(r$IRG_model), type = "o", pch = 16,
  #      ylab = "month", xlab = "", col = 3,
  #      ylim = c(0,12), main = paste(p$Location[1], "model"))
  # points(r$year, month(r$EVI_model), type ="o", pch = 16)
  # points(x = p$year, y = month(p$month), col = 2, pch = 16)
  # points(x = p$year, y = month(p$month), col = 2, pch = 16)
  
}

```
## months to phen
# Time to peak
## months
```{r}
IDs <- unique(Peaks$geeID)

Peaks$precip_spline <- ymd("1900-01-01")
Peaks$EVI1_model <- ymd("1900-01-01")
Peaks$IRG1_model <- ymd("1900-01-01")
Peaks$GRD1_model <- ymd("1900-01-01")
Peaks$EVI2_model <- ymd("1900-01-01")
Peaks$IRG2_model <- ymd("1900-01-01")
Peaks$GRD2_model <- ymd("1900-01-01")
Peaks$EVI_model <- ymd("1900-01-01")
Peaks$IRG_model <- ymd("1900-01-01")
Peaks$GRD_model <- ymd("1900-01-01")
Peaks$LOS_start <- ymd("1900-01-01")
Peaks$LOS_end <- ymd("1900-01-01")
Peaks$LOS <- NA
i = 1
for(i in 1:length(IDs)){
  pidx <- which(Peaks$geeID == IDs[i])
  p <- Peaks[pidx,]  
  r <- rs[rs$geeID == IDs[i],]
  j = 1
  for(j in 1:nrow(p)){
    
    # modis model
    Peaks$EVI_model[pidx[j]] <- round_date(
      ymd(r$EVI_model[which.min(abs(p$month[j] - as.Date(r$EVI_model)))]), 
      unit = "month")
    Peaks$IRG_model[pidx[j]] <- round_date(
      ymd(r$IRG_model[which.min(abs(p$month[j] - as.Date(r$IRG_model)))]), 
      unit = "month")
    Peaks$GRD_model[pidx[j]] <- round_date(
      ymd(r$GRD_model[which.min(abs(p$month[j] - as.Date(r$GRD_model)))]), 
      unit = "month")
    Peaks$EVI1_model[pidx[j]] <- round_date(
      ymd(r$EVI1_model[which.min(abs(p$month[j] - as.Date(r$EVI1_model)))]), 
      unit = "month")
    Peaks$IRG1_model[pidx[j]] <- round_date(
      ymd(r$IRG1_model[which.min(abs(p$month[j] - as.Date(r$IRG1_model)))]), 
      unit = "month")
    Peaks$GRD1_model[pidx[j]] <- round_date(
      ymd(r$GRD1_model[which.min(abs(p$month[j] - as.Date(r$GRD1_model)))]), 
      unit = "month")
    Peaks$EVI2_model[pidx[j]] <- round_date(
      ymd(r$EVI2_model[which.min(abs(p$month[j] - as.Date(r$EVI2_model)))]), 
      unit = "month")
    Peaks$IRG2_model[pidx[j]] <- round_date(
      ymd(r$IRG2_model[which.min(abs(p$month[j] - as.Date(r$IRG2_model)))]), 
      unit = "month")
    Peaks$GRD2_model[pidx[j]] <- round_date(
      ymd(r$GRD2_model[which.min(abs(p$month[j] - as.Date(r$GRD2_model)))]), 
      unit = "month")
    # Peaks[pidx[j],10:19]
  }
}

Peaks$months_to_precip_spline <- round(difftime(Peaks$month, 
                                                Peaks$precip_spline, 
                                                unit = "weeks")/4,0) %>% as.numeric
Peaks$months_to_EVI_model <- round(difftime(Peaks$month, 
                                            Peaks$EVI_model, 
                                            unit = "weeks")/4,0) %>% as.numeric
Peaks$months_to_IRG_model <- round(difftime(Peaks$month, 
                                            Peaks$IRG_model, 
                                            unit = "weeks")/4,0) %>% as.numeric
Peaks$months_to_GRD_model <- round(difftime(Peaks$month, 
                                            Peaks$GRD_model, 
                                            unit = "weeks")/4,0) %>% as.numeric
Peaks$months_to_EVI1_model <- round(difftime(Peaks$month, 
                                             Peaks$EVI1_model, 
                                             unit = "weeks")/4,0) %>% as.numeric
Peaks$months_to_IRG1_model <- round(difftime(Peaks$month, 
                                             Peaks$IRG1_model, 
                                             unit = "weeks")/4,0) %>% as.numeric
Peaks$months_to_GRD1_model <- round(difftime(Peaks$month, 
                                            Peaks$GRD1_model, 
                                            unit = "weeks")/4,0) %>% as.numeric
Peaks$months_to_EVI2_model <- round(difftime(Peaks$month, 
                                             Peaks$EVI2_model, 
                                             unit = "weeks")/4,0) %>% as.numeric
Peaks$months_to_IRG2_model <- round(difftime(Peaks$month, 
                                             Peaks$IRG2_model, 
                                             unit = "weeks")/4,0) %>% as.numeric
Peaks$months_to_GRD2_model <- round(difftime(Peaks$month, 
                                            Peaks$GRD2_model, 
                                            unit = "weeks")/4,0) %>% as.numeric


```

### plot
```{r}
library(data.table)
p <- as.data.table(Peaks)
p_mlt <- melt(p, id.vars = c("Location", "Country", "geeID", 
                             "Long", "Lat", "date", "ratio", 
                             "Count", "Observer"), 
              measure.vars = c("months_to_IRG_model","months_to_EVI_model",
                               "months_to_GRD_model",
                               "months_to_IRG1_model","months_to_EVI1_model",
                               "months_to_GRD1_model",
                               "months_to_IRG2_model","months_to_EVI2_model",
                               "months_to_GRD2_model",
                            "months_to_precip_spline"))
ggplot(p_mlt, aes(x = variable, y = value, fill = variable))+
  geom_hline(yintercept = 0)+
  geom_violin() + geom_boxplot(width=.1)+
  theme(legend.position = "none",
  axis.text.x = element_text(angle = 30, vjust = 1, hjust=1))+
  ylim(c(-10,7))
```

## prep data
```{r}
Peaks$abs_Lat <- abs(Peaks$Lat)
Peaks$log_colony_size <- log(Peaks$Count)
median(Peaks$months_to_IRG_model)
Peaks$abs_months_to_IRG_model <- abs(Peaks$months_to_IRG_model)
Peaks$abs_months_to_IRG1_model <- abs(Peaks$months_to_IRG1_model)

plot(Peaks$months_to_IRG_model, Peaks$months_to_IRG1_model)
hist(Peaks$months_to_IRG1_model)
```

## add lat long to rs
```{r}
rs$seasonality <- NA
IDs
for(i in 1:length(IDs)){
  rs$Lat[rs$geeID == IDs[i]] <- Peaks$Lat[Peaks$geeID == IDs[i]]
  rs$Long[rs$geeID == IDs[i]] <- Peaks$Long[Peaks$geeID == IDs[i]]
  # rs$seasonality[rs$geeID == IDs[i]] <- Peaks$seasonality[Peaks$geeID == IDs[i]][1]
}

rs$abs_Lat <- abs(rs$Lat)
Peaks$abs_months_to_IRG_model <- abs(Peaks$months_to_IRG_model)

```

## join peaks and rs
```{r}
library(geosphere)
rs$LOS <- rs$GRD_model - rs$IRG_model
rs$LOS1 <- rs$GRD1_model - rs$IRG1_model

rs$avg_LOS <- NA
IDs <- unique(rs$geeID)
i = 1
for(i in 1:length(IDs)){
  idx <- which(rs$geeID == IDs[i])
  SOS <- circ.mean.yday(yday(rs$IRG_model[idx]))
  EOS <- circ.mean.yday(yday(rs$GRD_model[idx]))
  if((EOS - SOS)>0) rs$avg_LOS[idx] <- EOS - SOS  
  if((EOS - SOS)<0) rs$avg_LOS[idx] <- SOS - EOS
}

Peaks$year <- year(Peaks$date)
for(i in 1:nrow(Peaks)){
  idx <- which(Peaks$geeID[i] == rs$geeID & Peaks$year[i] == rs$year)
  Peaks$amp[i] <-max(rs$amp[idx], na.rm = TRUE)
  # Peaks$seasonality[i] <- rs$seasonality[idx]
  Peaks$max_EVI[i] <- max(rs$max_EVI[idx], na.rm = TRUE)
  Peaks$min_EVI[i] <- max(rs$min_EVI[idx], na.rm = TRUE)
  Peaks$max_MEVI[i] <- max(rs$max_MEVI[idx], na.rm = TRUE)
  Peaks$min_MEVI[i] <- max(rs$min_MEVI[idx], na.rm = TRUE)
  Peaks$LOS[i] <-     max(rs$LOS[idx], na.rm = TRUE)
  Peaks$avg_LOS[i] <- max(rs$avg_LOS[idx], na.rm = TRUE)
  # Peaks$auc_EVI[i] <- max(rs$area_EVI[idx], na.rm = TRUE)
  Peaks$daylength[i]<-max(daylength(Peaks$Lat[i], 1:365))
}
Peaks$amp_EVI <- Peaks$max_EVI - Peaks$min_EVI
Peaks$amp_MEVI <- Peaks$max_MEVI - Peaks$min_MEVI

Peaks$log_colony_size <- log(Peaks$Count)
Peaks$abs_Lat <- abs(Peaks$Lat)

summary(Peaks)
Peaks[which(is.na(Peaks$min_EVI)),]
rs


```


```{r}

psych::pairs.panels(Peaks[,c("abs_months_to_IRG_model", "abs_Lat", "log_colony_size", 
                             "ratio", "max_EVI", "min_EVI",
                             "amp_EVI", "max_MEVI", "min_MEVI", 
                             "amp_MEVI",  "LOS", "avg_LOS")])
```

###### check for spatial autocorrelation
```{r}
# moran's I
library(tidyverse)
library(gridExtra)
library(NLMR)
library(DHARMa)

# unique_peaks <- Peaks %>% group_by(Lat, Long) %>% 
#   summarise(months_to_EVI = median(months_to_IRG_model),
#             season = median(seasonality))
# 
# m_non <- lm(months_to_EVI ~ season, unique_peaks)
# 
# sims <- simulateResiduals(m_non)
# # sims_rc <- recalculateResiduals(m_non)
# testSpatialAutocorrelation(sims, x = unique_peaks$Long, y = unique_peaks$Lat, plot = TRUE)
```

###### identify family for model
```{r}
library(fitdistrplus)
descdist(Peaks$months_to_IRG_model %>% abs, boot = 1000)
```


#### dredge

##### abs
```{r}
library(lme4)
library(MuMIn)
summary(Peaks)


Peaks[which(is.na(Peaks$min_EVI)),]

no_kasanka <-  Peaks[which(Peaks$geeID != 14),]
no_accra <- Peaks[which(Peaks$geeID != 10),]

Peaks$months_to_IRG_model %>% abs %>% 
  hist(breaks = 6, main = "", xlab = "Months to Peak IRG")
table(Peaks$abs_months_to_IRG_model)
m1 = glmer(abs(months_to_IRG_model)~
             scale(LOS)+ # correlated with max MEVI
             scale(avg_LOS)+
             scale(amp_MEVI)+
             scale(abs_Lat)+
             #scale(auc_EVI)+ # too correlated
             # scale(max_EVI)+
             #scale(min_EVI)+
             # scale(max_MEVI)+ # correlated with LOS
             scale(min_MEVI)+  # correlated with amp
             # scale(daylength)+ #correlated with amp and min EVI
             # scale(log(count))+
             scale(log_colony_size)+
             scale(ratio)+
             (1|geeID),
          family = poisson(link = "log"),
           data = Peaks, #no_kasanka,#
          na.action=na.fail)
summary(m1)
r.squaredGLMM(m1)
sims <- DHARMa::simulateResiduals(m1)
plot(sims)
dd = dredge(m1)
dd

# write.table(dd, file = "../output/dredge_absmonthstoIRG.txt")
```
### best
```{r}
m1_best = glmer(abs(months_to_IRG_model)~
             scale(abs_Lat)+
             scale(LOS)+
             # scale(max_EVI)+
             # scale(max_MEVI)+
             scale(log_colony_size)+
             
               # scale(min_EVI)+
               # scale(ratio)+
             (1|geeID),
          family = poisson(link = "log"),
           data = no_kasanka,#
          na.action=na.fail)
summary(m1_best)
r.squaredGLMM(m1_best)
sims <- DHARMa::simulateResiduals(m1_best)
plot(sims)
```
### interaction
```{r}
# m1i = glmer(abs(months_to_IRG_model)~
#              (scale(avg_LOS)+
#              #scale(amp_EVI)+
#              #scale(auc_EVI)+
#              scale(max_EVI)+
#              scale(min_EVI)+
#              # scale(daylength)+
#              # scale(log(count))+
#              scale(log_colony_size)+
#              scale(ratio))^2+
#              (1|geeID),
#           family = poisson(link = "log"),
#            data = Peaks, #no_kasanka,#
#           na.action=na.fail)
#           #REML=FALSE)
# summary(m1i)
# r.squaredGLMM(m1i)
# sims <- simulateResiduals(m1i)
# plot(sims)
# ddi = dredge(m1i)
# ddi
# ddi[1]
```

### Leave one out cross validation
```{r}
IDs <- unique(reg_colonies$geeID)
fits <- list()
r.sq <- list()

Peaks$predictIRG <- NA
i = 1

for(i in 1:length(IDs)){
  df <- Peaks[Peaks$geeID != IDs[i],]
  
  fits[[i]] <- lmer((months_to_IRG_model) ~ 
                     scale(LOS)+
                     scale(abs_Lat)+
                     scale(log_colony_size)+
                     # scale(ratio)+
                   (1|geeID), 
             data = df)
  summary(fits[[i]])
  r.sq[[i]] <-  r.squaredGLMM(fits[[i]])
  
  df_out <- Peaks[Peaks$geeID == IDs[i],]
  Peaks$predictIRG[which(Peaks$geeID == IDs[i])] <- 
    predict(fits[[i]], df_out, allow.new.levels = TRUE)
}


library(ggpmisc)
# https://stackoverflow.com/questions/7549694/add-regression-line-equation-and-r2-on-graph
my.formula <- y ~ x

ggplot(Peaks, aes(abs(months_to_IRG_model), predictIRG, col = Location))+
  geom_point()+
  geom_smooth(method = "lm", col = 1)+
  theme_classic()+
  stat_poly_eq(formula = my.formula, 
                aes(label = paste(..rr.label.., sep = "~~~")), 
                parse = TRUE)

r.sq_df <- sapply(r.sq, unlist)

plot(r.sq_df[2,], ylab = "R^2", xlab = "Dropped Colony",
     main = "Leave one Colony Out Model Performance")
abline(h = mean(r.sq_df[2,]))
```
## spatial glmm
```{r}
library(glmmfields)
library(ggplot2)
library(dplyr)
options(mc.cores = parallel::detectCores())

m_spatial <- glmmfields(as.numeric(months_to_IRG_model) ~
                          log(count) + amp + ratio + seasonality + (1|geeID),
  data = Peaks, 
  #family = Gamma(link = "log"),
  lat = "Lat", lon = "Long", 
  nknots = 12, iter = 2000, chains = 4,
  prior_intercept = student_t(50, 200, 70), 
  prior_beta = student_t(50, -30, 3),
  prior_sigma = half_t(50, 0, 3),
  prior_gp_theta = half_t(50, 0, 10),
  prior_gp_sigma = half_t(50, 0, 3),
  seed = 123 # passed to rstan::sampling()
)
m_spatial
```

### plot spatial glmm
```{r}
plot(m_spatial, type = "spatial-residual", link = TRUE) +
  geom_jitter(size = 3, width = 1, height = 1)
```


```{r}
plot(m_spatial, type = "residual-vs-fitted")
```


## gam
```{r}
library(mgcv)
#Peaks$daylength <- geosphere::daylength(Peaks$Lat, yday(Peaks$month))
Peaks$Location <- as.factor(Peaks$Location)
```

### normal
```{r}
g <- gam(abs(months_to_IRG_model)~
             # s(avg_LOS)+
             # s(LOS)+
             s(amp_EVI, k = 10)+ # R2 = 0.32
             # s(auc_EVI, k = 5)+
             s(max_EVI, k = 10)+
             # s(min_EVI, k = 5)+
             # s(daylength, k = 10)+ # r^2 = 0.37
             # s(log_colony_size, k = 10)+
             # s(ratio, k = 5)+
            # s(Long, Lat, k = 40, bs = "gp")+
            s(Location, bs = "re"),
            data = Peaks,
          method = "REML")
summary(g)
plot(g, scale = 0, shade = TRUE, scheme = 2)
gam.check(g, rep = 500)
```

#### spatial
```{r}
gs <- gam(abs(months_to_IRG_model)~
            s(max_EVI)+
            s(daylength)+
            s(Long, Lat, k = 10, bs = "tp")+
            s(Location, bs = "re"),
            data = Peaks,
          method = "REML")
summary(gs)
plot(gs, scale = 0, shade = TRUE, scheme = 2)
gam.check(gs, rep = 500)
AIC(g, gs)
```


### nb
```{r}
g_nb <- gam(abs(months_to_IRG_model)~
            s(max_EVI)+
            s(daylength)+
            #s(Long, Lat, k = 10, bs = "tp")+
            s(Location, bs = "re"),
            data = Peaks,
            family = nb,
          method = "REML")
summary(g_nb)
plot(g_nb, scale = 0, shade = TRUE, scheme = 2)
gam.check(g_nb, rep = 500)
```


#### spatial
```{r}
g_nbs <- gam(abs(months_to_IRG_model)~
            s(max_EVI)+
            s(daylength)+
            s(Long, Lat, k = 10, bs = "tp")+
            s(Location, bs = "re"),
            data = Peaks,
            family = nb,
          method = "REML")
summary(g_nbs)
plot(g_nbs, scale = 0, shade = TRUE, scheme = 2)
gam.check(g_nbs, rep = 500)
```

### tw
```{r}
g_tw <- gam(abs(months_to_IRG_model)~
            s(max_EVI)+
            s(daylength)+
            s(Location, bs = "re"),
            data = Peaks,
            family = tw,
          method = "REML")
summary(g_tw)
plot(g_tw, scale = 0, shade = TRUE, scheme = 2)
gam.check(g_tw, rep = 500)

AIC(g, g_nb, g_tw)
```
#### spatial
```{r}
g_tws <- gam(abs(months_to_IRG_model)~
            s(max_EVI)+
            s(daylength)+
            
              s(Long, Lat, k = 10, bs = "tp")+
            s(Location, bs = "re"),
            data = Peaks,
            family = tw,
          method = "REML")
summary(g_tws)
plot(g_tws)
gam.check(g_tws, rep = 500)

aics <- AIC(g, gs, g_nb, g_nbs, g_tw, g_tws)
aics[order(aics$AIC),]
```

### ziP
```{r}
g_ziP <- gam(abs(months_to_IRG_model)~
            s(max_EVI)+
            s(daylength)+
            s(Location, bs = "re"),
            data = Peaks,
            family = ziP,
          method = "REML")
summary(g_ziP)
plot(g_ziP, scale = 0, shade = TRUE, scheme = 2)
gam.check(g_ziP, rep = 5000)

AIC(g, g_nb, g_tw, g_ziP)
```
#### spatial
```{r}
g_ziPs <- gam(abs(months_to_IRG_model)~
            s(max_EVI, k = 5)+
            # s(irg, k = 5)+
            # s(precip)+
            s(daylength)+
            # s(month)+
            # s(year, k = 20)+
            s(Long, Lat, k = 10, bs = "tp"),
            #s(Location, bs = "re"),
            data = Peaks,
            family = ziP,
          method = "REML")
summary(g_ziPs)
plot(g_ziPs, scale = 0, shade = TRUE, scheme = 2)
gam.check(g_ziPs, rep = 500)

AIC(g, g_nb, g_tw, g_ziP, gs, g_nbs, g_tws, g_ziPs)
```

